Online Verification of Commutativity

### 2020-09-10 acmPaper/paper.tex
Little tweaks in 4

We present a polynomial time solution to the \textsc{verification set problem} .
As in the online baseline algorithm, we do not concern ourselves with parallel pairs where neither or both paths pass through the new edge.
The key observation allowing us to improve on the online baseline is a result of Theorem~\ref{reductionRule} (which will be expanded on later): for a given source and sink pair, only a single parallel pair needs to be verified.
Theorem~\ref{verifyingSet} shows that, should our selected set of pairs and cycles passing through the new edge be verified commutative, the entire diagram must commute.
Algorithm \ref{algo_online_polynomial} uses this strategy of identifying a parallel pair with exactly one path through the edge for each (source, sink) pair.

\begin{algorithm}
\DontPrintSemicolon
\KwData{existing graph, new edge.}
\KwResult{Set of parallel pairs to verify.}
\SetKwProg{try}{try}{:}{}
\SetKwProg{catch}{catch}{:}{end}
Graph existingGraph;
Edge newEdge;

@@

The try block is executed at most $O(|V|^2)$ times, which is also the bound on the number of pairs verified.
%
This bound is asymptotically tight, as can be seen in the case where the graph contains $2N$ nodes along $S$ and $T$.
We imagine dividing the nodes into two groups of $N$ nodes each. Every node in group 1 has a forward node to every node in group 2 as well as to $S$. $T$ has a forward edge to every node in group 2. In this diagram, on the addition of edge $(S, T)$, $N^2$ paths need to be verified which is polynomial in the total number of nodes, $2N+2$.

@@

\begin{theorem}
\label{reductionRule}
If parallel paths $g_2 = f_2; n; h_2$ then it must be that $g_1 = f_1; n; h_1$.
\end{theorem} 
% monomorphisms
\begin{proof}
We use the fact that $f_1$=$l; f_2$ and $h_1$=$h_2; m$.
\[g_2 = f_2; n; h_2 \Rightarrow l; g_2 = l; f_2; n ; h_2 \]
\[\Rightarrow l ; g_2 ; m = l ; f_2 ; n ; h_2 ; m \Rightarrow g_1 = f_1 ; n ; h_1\]

The proof is not affected if any of the paths used are the identity, e.g., if $f_1$ is the identity and $S$ and $P_1$ are the same node.
\end{proof}

@@

 Given that a set of path pairs are equal, suppose that one attempts to derive the proposition that a different parallel pair of paths is equal with a step-by-step application of inference rules.
 Under the assumption that edges are generic functions, and no other information is available, $F$ is a semi-group.
 The only inference rules allowed are composition (given that $f_1=f_2$, it must be that $g ; f_1 = g ; f_2$) and replacement of one path by a different, equal path (given $f_1=f_2$ and $g ; f_1 = h ; f_1 $, it must be true that $g ; f_1 = h ; f_2$). Any permutation of the repeated application of these two rules results in the ``reduction rule'' already described; it is therefore the only rule that can be used to reduce the set of path pairs to check.

>>>>

We present a polynomial time solution to the \textsc{verification set problem}.
As in the online baseline algorithm, we do not concern ourselves with parallel pairs where neither or both paths pass through the new edge.
The key observation allowing us to improve on the online baseline is a result of Theorem~\ref{reductionRule} (which we expand on later): for a given source and sink pair, only a single parallel pair needs to be verified.
Theorem~\ref{verifyingSet} shows that, should our selected set of pairs and cycles passing through the new edge be verified commutative, the entire diagram must commute.
\ref{algo_online_polynomial} uses this strategy of identifying a parallel pair with exactly one path through the edge for each (source, sink) pair.

\begin{algorithm}[t]
\DontPrintSemicolon
\KwData{existing graph, new edge.}
\KwResult{Set of parallel pairs to verify.}
\SetKwProg{try}{try}{:}{}
\SetKwProg{catch}{catch}{:}{end}
Graph existingGraph;
Edge newEdge;

@@

The \code{try} block is executed at most $O(|V|^2)$ times, which is also the bound on the number of pairs verified.
%
This bound is asymptotically tight, as can be seen in the case where the graph contains $2N$ nodes along $S$ and $T$.
Imagine dividing the nodes into two groups of $N$ nodes each. Every node in group 1 has a forward edge to every node in group 2 and to $S$. $T$ has a forward edge to every node in group 2. In this diagram, when adding edge $(S, T)$, $N^2$ paths need to be verified which is polynomial in the total number of nodes, $2N+2$.

@@

\begin{theorem}
\label{reductionRule}
If parallel paths $g_2 = f_2; n; h_2$ then it must be that $g_1 = f_1; n; h_1$.
\end{theorem} 
% monomorphisms
\begin{proof}
We use the fact that $f_1$=$l; f_2$ and $h_1$=$h_2; m$.
\[g_2 = f_2; n; h_2 \Rightarrow l; g_2 = l; f_2; n ; h_2 \]
\[\Rightarrow l ; g_2 ; m = l ; f_2 ; n ; h_2 ; m \Rightarrow g_1 = f_1 ; n ; h_1\]
%
The proof holds if any of the paths used are the identity, e.g., if $f_1$ is the identity so $S$ and $P_1$ are the same node.
\end{proof}

@@

Given that a set of path pairs are equal, suppose we attempt to derive the proposition that a different parallel pair of paths is equal with a step-by-step application of inference rules.
 Under the assumption that edges are generic functions, and no other information is available, $F$ is a semi-group.
 The only inference rules allowed are composition (given that $f_1=f_2$, it must be that $g ; f_1 = g ; f_2$) and replacement of one path by a different, equal path (given $f_1=f_2$ and $g ; f_1 = h ; f_1 $, it must be true that $g ; f_1 = h ; f_2$). Any permutation of the repeated application of these two rules results in the ``reduction rule'' already described; it is therefore the only rule that can be used to reduce the set of path pairs to check.
